Correlation tests

### 2025-03-19 analysis/cor_tables.qmd
Testing correlation

---
title: "Correlation tests"
format: 
    gfm:
      echo: true
      message: false
      warning: false
      code-fold: true
prefer-html: true
---

@@

The aim of this document is to determine the correlation between coefficients for several hybrid open access indicators between hoaddata, Web of Science and Scopus.

@@

  # Create a nicer version of variable names for display
  nice_names <- gsub("_", " ", vars)
  nice_names <- stringr::str_to_title(nice_names)

  # Set row and column names
  rownames(corr_df) <- nice_names
  colnames(corr_df) <- nice_names

  return(list(df = corr_df, title = paste("Spearman Correlation Matrix -", group_name)))
}

#' Display correlation matrix using Pandoc table with row names
#'
#' @param corr_list A list containing the correlation data frame and title
#' @return A character string containing the Pandoc table
create_pandoc_table <- function(corr_list) {
  footnote_text <- c("* p < 0.05", "** p < 0.01", "*** p < 0.001")

@@

  # Add title
  table_title <- paste("#", corr_list$title, "\n\n")

@@

### Article volume 2019-23

@@

>>>>

---
title: "Correlation tests"
format: 
    gfm:
      echo: true
      message: false
      warning: false
      code-fold: true
bibliography: references.bib
---

@@

The aim of this document is to determine the correlation coefficients for several hybrid open access indicators between hoaddata, Web of Science and Scopus, focusing on country affiliations between 2019 and 2023.
Related research have compared open scholarly data sources and propietory databases at the country-level using Spearman's rank correlation coefficient [@alperin2024analysissuitabilityopenalexbibliometric] and Kendal tau correlation [@Akbaritabar_2024], showing high level of correlations.

@@

  return(list(df = corr_df, title = paste0(group_name, "--Spearman Correlation Matrix")))
}

#' Display correlation matrix using Pandoc table with row names
#'
#' @param corr_list A list containing the correlation data frame and title
#' @return A character string containing the Pandoc table
create_pandoc_table <- function(corr_list) {
  footnote_text <- c("\\* p < 0.05", "** p < 0.01", "*** p < 0.001")

@@

  # Add title
  table_title <- paste("###", corr_list$title, "\n\n")

@@

@@

oa_articles <- create_correlation_table(my_df_ind, vars =  c("hoad_oa_articles", "scp_oa_first_corresponding_author_articles", "scp_oa_corresponding_author_articles",
            "wos_oa_first_author_articles", "wos_oa_corresponding_author_articles"), "Open Access volume")
cat(create_pandoc_table(oa_articles))
```

```{r}
#| results: asis
oa_prop_ind <- my_df_ind |>
  # As in Figure
  mutate(
    hoad_oa_prop = hoad_oa_articles / hoad_articles,
    scp_first_oa_prop = scp_oa_first_author_articles / scp_first_author_articles,
    scp_corresponding_oa_prop = scp_oa_corresponding_author_articles / scp_corresponding_author_articles,
    wos_first_oa_prop = wos_oa_first_author_articles / wos_first_author_articles,
    wos_corresponding_oa_prop = wos_oa_corresponding_author_articles / wos_corresponding_author_articles
  ) |>
  select(country_code, contains("prop")) |>
  create_correlation_table(vars =  c("hoad_oa_prop", "scp_first_oa_prop", 
            "scp_corresponding_oa_prop", "wos_first_oa_prop", "wos_corresponding_oa_prop"), "Open Access Uptake")
cat(create_pandoc_table(oa_prop_ind))
```

```{r}
#| results: asis
oa_prop_ind <- my_df_ind |>
  # As in Figure
  filter(!is.na(country_code), hoad_articles > 10000) |>
  mutate(
    hoad_oa_prop = hoad_oa_articles / hoad_articles,
    scp_first_oa_prop = scp_oa_first_author_articles / scp_first_author_articles,
    scp_corresponding_oa_prop = scp_oa_corresponding_author_articles / scp_corresponding_author_articles,
    wos_first_oa_prop = wos_oa_first_author_articles / wos_first_author_articles,
    wos_corresponding_oa_prop = wos_oa_corresponding_author_articles / wos_corresponding_author_articles
  ) |>
  select(country_code, contains("prop")) |>
  create_correlation_table(vars =  c("hoad_oa_prop", "scp_first_oa_prop", 
            "scp_corresponding_oa_prop", "wos_first_oa_prop", "wos_corresponding_oa_prop"), "Open Access Uptake (min 1.000 articles)")
cat(create_pandoc_table(oa_prop_ind))
```

ta_oa_articles <- create_correlation_table(my_df_ind, vars = c("hoad_ta_oa_articles", "scp_ta_oa_first_corresponding_author_articles", "scp_ta_oa_corresponding_author_articles",
            "wos_ta_oa_first_author_articles", "wos_ta_oa_corresponding_author_articles"), "Open Access volume via Transformative Agreements")
cat(create_pandoc_table(ta_oa_articles))
```

```{r}
#| results: asis
ta_oa_articles <- my_df_ind |>
    filter(!is.na(country_code), hoad_articles > 10000) |>
  create_correlation_table(vars = c("hoad_ta_oa_articles", "scp_ta_oa_first_corresponding_author_articles", "scp_ta_oa_corresponding_author_articles",
            "wos_ta_oa_first_author_articles", "wos_ta_oa_corresponding_author_articles"), "Open Access volume via Transformative Agreements (min 1,000 articles)")
cat(create_pandoc_table(ta_oa_articles))
```

ta_oa_prop <- my_df_ind |>
 #filter(hoad_ta_oa_articles > 1000) |>
  mutate(
    hoad_oa_prop = hoad_ta_oa_articles / hoad_oa_articles,
    scp_first_oa_prop = scp_ta_oa_first_author_articles / scp_oa_first_author_articles,
    scp_corresponding_oa_prop = scp_ta_oa_corresponding_author_articles / scp_oa_corresponding_author_articles,
    wos_first_oa_prop = wos_ta_oa_first_author_articles / wos_oa_first_author_articles,
    wos_corresponding_oa_prop = wos_ta_oa_corresponding_author_articles / wos_oa_corresponding_author_articles
  ) |>
  select(country_code, contains("prop")) |>
  create_correlation_table(vars =  c("hoad_oa_prop", "scp_first_oa_prop", 
            "scp_corresponding_oa_prop", "wos_first_oa_prop", "wos_corresponding_oa_prop"), "% of Hybrid OA via Agreement")
cat(create_pandoc_table(ta_oa_prop))
```

ta_oa_prop <- my_df_ind |>
 filter(hoad_ta_oa_articles > 1000) |>
  mutate(
    hoad_oa_prop = hoad_ta_oa_articles / hoad_oa_articles,
    scp_first_oa_prop = scp_ta_oa_first_author_articles / scp_oa_first_author_articles,
    scp_corresponding_oa_prop = scp_ta_oa_corresponding_author_articles / scp_oa_corresponding_author_articles,
    wos_first_oa_prop = wos_ta_oa_first_author_articles / wos_oa_first_author_articles,
    wos_corresponding_oa_prop = wos_ta_oa_corresponding_author_articles / wos_oa_corresponding_author_articles
  ) |>
  select(country_code, contains("prop")) |>
  create_correlation_table(vars =  c("hoad_oa_prop", "scp_first_oa_prop", 
            "scp_corresponding_oa_prop", "wos_first_oa_prop", "wos_corresponding_oa_prop"), "% of Hybrid OA via Agreement (min 1.000 OA articles)")
cat(create_pandoc_table(ta_oa_prop))
```

# References
